Supplemental Information A: Full Annotated Analysis and Reproducible Code

### 2022-06-27 Supplemental-Material-A-Annotated-Analysis.Rmd
MI, three-way ANOVA, Plots added + clean up

---
title: "Supplemental Information A: Full Annotated Analysis and Reproducible Code"
subtitle: "Supplemental Material for 'Migration Trajectories [working title]'"
author:
- Jannis Kreienkamp^1,3^
- Kai Epstude^1,3^
- Laura F. Bringmann^1,3^
- Maximilian Agostini^1,3^
- Peter de Jonge^1,3^
- Rei Tendeiro-Monden^2,3^
- ^1^University of Groningen, Department of Psychology
- ^2^Hiroshima University, Graduate School of Advanced Science and Engineering
- ^3^Author order TBD [currently in first name alphabetical order]
- "Author Information:"
- "Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).  E-mail: j.kreienkamp@rug.nl"
- 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
- 'The data repository for this manuscript is available at <a href="https://osf.io/TBA" target="_blank">osf.io/TBA</a>'
- 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    fig_caption: yes
    md_extensions: +footnotes
    code_folding: hide
    mathjax: default
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: false
    css: style.css
    includes:
      in_header: "_includes/head-custom-rmd.html" 
editor_options:
  chunk_output_type: console
#bibliography: references.bib
csl: apa.csl
header-includes:
   - \usepackage{amsmath, nccmath}
---

@@

df %>%
  #filter(OutgroupInteraction == "Yes") %>%
  select(
    InteractionContextEqualStatus,
    KeyNeedShared,
    InteractionContextCooperative,
    InteractionContextvoluntary
  ) %>%
  psych::describe(., skew=F,ranges=T) %>%
  as.data.frame() %>%
  select(-vars) %>%
  kable(., caption = "Descriptives of Allport's Condition items") %>% 
  kable_styling("hover", full_width = F, latex_options = "hold_position")

@@

sjPlot::tab_itemscale(iaWorkerAllport)

@@

sjPlot::tab_itemscale(iaWorkerAllport, factor.groups)

ltm::cronbach.alpha(na.omit(iaWorkerAllport), CI = TRUE)

@@

horst::nestedAlpha(item.level.1 = "value",
                   level.2      = "TIDnum",
                   level.3      = "PID",
                   data         = data)
rm(data)

@@

## Data Availability and Sample Selection  {.tabset}

@@

@@

### Timescale ML test

```{r }
dtMain3mpca <- dtMain3mpca %>%
  mutate(
    TIDdate = as.numeric(factor(date)),
    TIDweek = as.numeric(factor(week))
  )

test <-
  lme4::lmer(KeyNeedFulfillment ~ 1 + (1 + TIDnum + TIDdate + TIDweek | PID),
    data = dtMain3mpca
  ) # use optim if it does not converge

test2 <-
  lme(
    KeyNeedFulfillment ~ 1,
    random = ~ 1 + TIDnum + TIDdate + TIDweek | PID,
    data = dtMain3mpca %>% filter(!is.na(KeyNeedFulfillment)),
    control = list(opt = "nlmimb")
  ) # use optim if it does not converge

summary(test2)
#summ(test, digits = 3, center = FALSE)
anova(test2)

@@

@@

@@

varNamInMiMainPsbl <- c(
  paste(varNamInMiMain, rep(".pre", length(varNamInMiMain)), sep = ""),
  varNamInMiMain,
  paste(varNamInMiMain, rep(".post", length(varNamInMiMain)), sep = "")
)

@@

dtMiMain <- rbind(
  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% mutate(across(!TID &
                                                                                             !study, as.numeric)),
  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(date = as.Date(gsub(" .*", "", TID)),
         week = strftime(date, format = "%Y-W%V")) %>%
  select(ID,
         study,
         date,
         week,
         everything(),
         -PID,
         -starts_with("ResponseId")) %>%
  as.data.frame %>%
  select_if( ~ sum(!is.na(.)) > 1) # only include variables that have any data (i.e., not all NA)

MiMain <-
  amelia(
    dtMiMain,
    m = 1, # 20
    ts = "TIDnum",
    cs = "ID",
    idvars = c("TID", "date", "week", "study"), #"PID", 
    noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
    parallel = "multicore",
    ncpus = parallel::detectCores(),
    p2s = 1
  )
save(MiMain, file = "data/imputed/MainImputed.RData")
```

@@

@@

varNamInMiS1Psbl <- c(
  paste(varNamInMiS1, rep(".pre", length(varNamInMiS1)), sep = ""),
  varNamInMiS1,
  paste(varNamInMiS1, rep(".post", length(varNamInMiS1)), sep = "")
)

@@

dtMiS1 <- dtS1Red %>%
  select(
    PID,
    TID,
    TIDnum,
    date,
    week,
    any_of(varNamInMiS1Psbl),
    -time,
    -starts_with("ResponseId")
  ) %>%
  mutate(
    date = as.Date(gsub(" .*", "", TID)),
    week = strftime(date, format = "%Y-W%V")
    ) %>%
  as.data.frame %>%
  select_if(~sum(!is.na(.)) > 1) #%>% # only include variables that have any data (i.e., not all NA)

  which(!is.numeric(dtMiS1) | sapply(dtMiS1, sd, na.rm=TRUE)>0)

  select_if(~sd(., na.rm = TRUE) > 0) # only include variables that vary (i.e., sd != 0)

@@

MiS1 <-
  amelia(
    dtMiS1,
    m = 2,
    ts = "TIDnum",
    cs = "PID",
    idvars =  c("TID", "date", "week"),
    noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
    parallel = "multicore",
    ncpus = parallel::detectCores(),
    p2s = 1
  )
write.amelia(MiS1, file.stem = "data/imputed/S1Imputed.Amelia")
```

@@

@@

@@

@@

@@

# test <- PCAnorm(data = dtPcaS123,
#           pid = "ID",
#           tid = "TIDnum",
#           selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>% 
#   select(
#     ID,
#     PID,
#     TIDnum,
#     study,
#     ends_with("_gmc")
#   ) %>%
#   data.frame %>%
#   melt(
#     .,
#     id=c("ID", "PID", "TIDnum", "study")
#   ) %>%
#   mutate(
#     variable = paste(gsub("_gmc", "", variable), TIDnum, sep = "_")
#   ) %>%
#   #select(-TIDnum) %>% 
#   arrange(variable, ID) %>%
#   mutate(
#     idVar = paste(ID, variable, sep = "_")
#   ) 
# plyr::join(test, data.frame(table(idVar = test$idVar)))

test <- PCAnorm(data = dtPcaS123,
          pid = "ID",
          tid = "TIDnum",
          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>%
  select(
    ID,
    PID,
    TIDnum,
    study,
    ends_with("_gmc")
  ) %>%
  data.frame %>% 
  melt(
    .,
    id=c("ID", "PID", "TIDnum", "study")
  ) %>%
  mutate(
    variable = paste(gsub("_gmc", "", variable), stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
  ) %>%
  select(-TIDnum) %>%
  arrange(variable, ID) %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  replace(is.na(.), 0) %>%
  select(-ID, -PID, -study)
  #reshape(., idvar = "ID", timevar = "variable", direction = "wide") %>% View
  #pivot_wider(names_from = variable, values_from = value)

test2 <- dtPcaS123 %>%
  select(
    -TID,
    -date,
    -week
  ) %>%
  data.frame %>% 
  melt(
    .,
    id=c("ID", "PID", "TIDnum", "study")
  ) %>%
  mutate(
    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
  ) %>%
  select(-TIDnum) %>%
  arrange(variable, ID) %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  replace(is.na(.), 0) %>%
  select(-ID, -PID, -study)

@@

```{r threeWayAnova}
library(ThreeWay)
threeWayAno <- threewayanova(
  Y = test2,
  n = nrow(test2), # Nr. ppt.s (A)
  m = length(unique(dtPcaS123$TIDnum)), # Nr. time points (B)
  p = sum(!varNamPcaS123 %in% idVars) # Nr. variables (C)
)

@@

# **Plots per Item** {.tabset}

@@

@@

### Studies 1-3 (Means)

```{r plotAllMean}
varClS123 <- dtPcaS123 %>%
  PCAnorm(data = .,
          pid = "PID",
          tid = "TIDnum",
          selection = names(.)[!names(.) %in% idVars]) %>%
  select(
    ends_with("_gmc")
  ) %>%
  data.frame %>%
  summarise_all(
    sd, na.rm = TRUE
  ) %>%
  t %>% 
  as.data.frame %>%
  kmeans(., centers = 2)

PCAnorm(data = dtPcaS123,
          pid = "ID",
          tid = "TIDnum",
          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>% 
  select(
    ID,
    TIDnum,
    ends_with("_gmc")
  ) %>%
  data.frame %>%
  melt(
    .,
    id=c("ID","TIDnum")
  ) %>%
  mutate(
    variable = gsub("_gmc", "", variable)
  ) %>%
  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
  #geom_jitter() +
  stat_summary(fun=mean, geom="line") +
  labs(
    title = "Studies 1-3 Means over time [bi-daily]",
    y = "Grand Mean Centered Average",
    x = "Variable",
    color = "Timepoint",
    group = "Timepoint"
  ) +
  #coord_flip() +
  theme_Publication() +
  theme(
    axis.text.x = element_text(angle = 45, hjust=1),
    legend.key.size = unit(1, 'cm'),
    legend.key.height = unit(0.4, 'cm'),
  )

@@

### Study 1 (Means)

@@

We begin by plotting the item means of all variables. We plot the means of the individual variables for each measurement occasion, summarizing the responses of all participants at any given time point. The displayed time axis is the measurement index rather than the actual date to ensure comparability of the changes between each measurement. We plot the means for bi-daily, daily, and weekly timescales.

@@

### Study 2 (Means)

@@

### Study 3 (Means)

@@

<font size="1"> This is my text number 1</font>

<font size="2"> This is my text number 2</font>

<font size="3"> This is my text number 3</font>

<font size="4"> This is my text number 4</font>

<font size="5"> This is my text number 5</font>

<font size="6"> This is my text number 6</font>

@@

### Studies 1-3 (SDs)

@@

### Study 1 (SDs)

@@

### Study 2 (SDs)

@@

### Study 3 (SDs)

@@

<!-- # **References**   -->

>>>>

---
title: "Supplemental Information A: Full Annotated Analysis and Reproducible Code"
subtitle: "Supplemental Material for 'Migration Trajectories [working title]'"
author:
- Jannis Kreienkamp^1,3^
- Kai Epstude^1,3^
- Laura F. Bringmann^1,3^
- Maximilian Agostini^1,3^
- Peter de Jonge^1,3^
- Rei Monden^2,3^
- ^1^University of Groningen, Department of Psychology
- ^2^Hiroshima University, Graduate School of Advanced Science and Engineering
- ^3^Author order TBD [currently in first name alphabetical order]
- "Author Information:"
- "Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).  E-mail: j.kreienkamp@rug.nl"
- 'The main manuscript is available at <a href="https://www.doi.org/ToBePublished" target="_blank">doi.org/ToBePublished</a>'
- 'The data repository for this manuscript is available at <a href="https://osf.io/TBA" target="_blank">osf.io/TBA</a>'
- 'The GitHub repository for this manuscript is available at <a href="https://github.com/maskedForPeerReview" target="_blank">github.com/maskedForPeerReview</a>'
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    fig_caption: yes
    md_extensions: +footnotes
    code_folding: hide
    mathjax: default
    theme: yeti
    toc: yes
    toc_float: yes
    number_sections: false
    css: style.css
    includes:
      in_header: "_includes/head-custom-rmd.html" 
editor_options:
  chunk_output_type: console
bibliography: references.bib
csl: apa.csl
header-includes:
   - \usepackage{amsmath, nccmath}
---

@@

AlportDescr <- df %>%
  #filter(OutgroupInteraction == "Yes") %>%
  select(
    InteractionContextEqualStatus,
    KeyNeedShared,
    InteractionContextCooperative,
    InteractionContextvoluntary
  ) %>%
  psych::describe(., skew=F,ranges=T) %>%
  as.data.frame() %>%
  select(-vars) %>%
  kable(., caption = "Descriptives of Allport's Condition items") %>% 
  kable_styling("hover", full_width = F, latex_options = "hold_position")

@@

itemScaleAllport01 <- sjPlot::tab_itemscale(iaWorkerAllport)

@@

sjPlot::tab_itemscale(iaWorkerAllport, factor.groups, show.kurtosis = TRUE)

AllportAlpha <- ltm::cronbach.alpha(na.omit(iaWorkerAllport), CI = TRUE)

@@

AllportNestedAlpha <- horst::nestedAlpha(item.level.1 = "value",
                   level.2      = "TIDnum",
                   level.3      = "PID",
                   data         = data)
rm(data)

@@

## Data Availability and Sample Selection {.tabset .tabset-fade}

@@

For our sample selection we address both the selection of time points and participants. Given that multiple imputation procedures even work well with large proportions of missing data [assuming missingness at random; @Madley-Dowd2019], we decided on a general criterion of less than 45% missingness to balance sample size retention and bias in the multiple imputation model. Thus, we then select the time points for which we have less than 45% missingness and select participants who have less than 45% missingness across the selected time range.

To approximate an ideal sample selection that maximized the number of participants and timepoints that fit the 55% data availability criterion. We select sequentially remove the rows or columns with the lowest data availability rate untill all rows and columns have more than 55% data available (i.e., less than 45% missingness).

<div class="alert alert-info alert-dismissible fade in" role="alert">
  <a class="close" data-dismiss="alert" aria-label="close">&times;</a>
  <i class="fas fa-info-circle"></i> <b>Note: </b><br/>
  This needs to be adjusted current missingness criterion set to more conservative 33%.
</div>

@@

We aggregate the key variables over time to archive a reasonably interpretable number of time points and remove a first proportion of missing data. Given that little data is available on the meaningful time scales of the selected psychological variables, we chose to determine the appropriate time scales using variance decomposition [e.g., see @Ram2014]. This is to say that we create multi-level unconditional means models (without predictors) that include possible nested time scales as levels. We chose to select time scales that align with common human cycles. We thus compare the variances of bi-daily, daily, and weekly aggregations. Additional aggregations of two weeks or the full four weeks might be possible but would most likely reduce the variance too much for any meaningul further reduction during the 3MPCA.
We then chose the time scales that have the most variance.

```{r keyNeedTest}

# WILL USE IMPUTED DATA AND FOR-LOOP ACROSS ALL RELEVANT DVs
# PLOT EFFECT SIZES

dtMain3mpca <- dtMain3mpca %>%
  group_by(study) %>%
  mutate(
    TIDdate = as.numeric(factor(date)),
    TIDweek = as.numeric(factor(week))
  ) %>%
  ungroup

# Not sure whether TIDs as predictors or empty model?
keyNeedMlTest01 <-
  lme4::lmer(KeyNeedFulfillment ~ 1 + TIDnum + TIDdate + TIDweek + (1 + TIDnum + TIDdate + TIDweek | ID),
    data = dtMain3mpca
  )

keyNeedMlTest02 <-
  lme(
    KeyNeedFulfillment ~ 1 + TIDnum + TIDdate + TIDweek,
    random = ~ 1 + TIDnum + TIDdate + TIDweek | ID,
    data = dtMain3mpca %>% filter(!is.na(KeyNeedFulfillment)),
    control = list(opt = "nlmimb")
  ) # use optim if it does not converge

# summary(keyNeedMlTest02)
# summ(keyNeedMlTest01, digits = 3, center = FALSE)
# anova(keyNeedMlTest02)

@@

We ultimately aim to create 20 imputed datasets to perform the analyses on. As outlined by @Monden2015, we will use the key variables themself as well as auxiliary variables to impute missing values. Auxiliary variables are any pre-, post-, or daily dairy variables other than the main variables. If there are any modeling issues we will limit the auxiliary variables to all variables that that are significantly ($p$ < .01) and meaningfully ($r$ > .3) correlated with (1) the key variables or (2) missingness on the key variables.

We create the imputed data set for each of the analyses. This includes the main analyses with across all three studies, using the set of common variables, as well as follow-up analyses for study-specific, interaction-specific and time scale specific datasets.

@@

We begin by preparing the data for the main analysis. Here, we:

1. Select the variables for all relevant analysis (i.e., imputation, 3MPCA, and validation) across all three studies.
2. Create the combined dataset in the correct format (long format with all missed time points as NAs).
3. Impute the data and save the imputed datasets for later recall.

@@

varNamMiMain <- c(varNamAdd, varNamInMiMain[!varNamInMiMain %in% varNamOut])

varNamInMiMainPsbl <- c(
  paste(varNamMiMain, rep(".pre", length(varNamInMiMain)), sep = ""),
  varNamMiMain,
  paste(varNamMiMain, rep(".post", length(varNamInMiMain)), sep = "")
)

@@

dtMiMain <- rbind(
  dtS1Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S1") %>% mutate(across(!TID &
                                                                                             !study, as.numeric)),
  dtS2Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S2"),
  dtS3Red %>% select(any_of(varNamInMiMainRed)) %>% mutate(study = "S3")
) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(
    date = as.Date(gsub(" .*", "", TID)),
    week = strftime(date, format = "%Y-W%V"),
    roomate.pre = as.numeric(roomate.pre),
    Reason.pre = as.numeric(Reason.pre),
    association.pre = as.numeric(association.pre)
  ) %>%
  select(ID,
         study,
         date,
         week,
         everything(),
         -PID,
         -starts_with("ResponseId")) %>%
  as.data.frame %>%
  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  complete(., ID, TIDnum) %>% # add na if id and TID combination is missing
  group_by(ID) %>%
  mutate(across(ends_with(c(".pre", ".post")),  ~ replace_na(.x, mean(.x, na.rm = TRUE)))) %>%
  mutate(study = replace_na(study, calc_mode(study))) %>%
  ungroup %>%
  group_by(study, TIDnum) %>%
  mutate(across(c("date", "week", "TID"),  ~ replace_na(.x, calc_mode(.x)))) %>%
  ungroup %>% 
  mutate(across(ends_with(c(".pre", ".post")), ~ ifelse(is.nan(.), NA, .))) %>%
  as.data.frame

# MiMain <-
#   amelia(
#     dtMiMain,
#     m = 1, # 20
#     ts = "TIDnum",
#     cs = "ID",
#     idvars = c("TID", "date", "week", "study"), #"PID",
#     #lags = names(dtMiMain)[!names(dtMiMain) %in% idVars],
#     #lags = names(dtMiMain)[names(dtMiMain) %in%varNamInMiMainRed],
#     #noms = names(dtMiMain)[grepl(paste(varNamMiMainNoms, collapse = "|^"), names(dtMiMain))],
#     parallel = "multicore",
#     ncpus = parallel::detectCores(),
#     p2s = 1
#   )
# save(MiMain, file = "data/imputed/MainImputed.RData")
load("data/imputed/MainImputed.RData")
```

@@

To be done once initial tests approved.

@@

varNamMiS1 <- c(varNamAdd, varNamInMiS1[!varNamInMiS1 %in% varNamOut])

varNamInMiS1Psbl <- c(
  paste(varNamMiS1, rep(".pre", length(varNamMiS1)), sep = ""),
  varNamMiS1,
  paste(varNamMiS1, rep(".post", length(varNamMiS1)), sep = "")
)

@@

# TO BE ADAPTED:
dtMiS1 <- dtS1Red %>%
  mutate(
    date = as.Date(gsub(" .*", "", TID)),
    week = strftime(date, format = "%Y-W%V")
    ) %>%
  select(
    PID,
    TID,
    TIDnum,
    date,
    week,
    any_of(varNamInMiS1Psbl),
    -time,
    -starts_with("ResponseId")
  ) %>%
  as.data.frame %>%
  mutate_at(vars(-c("TID", "date", "week")), as.numeric) %>%
  select_if(~sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  select(PID,
         TID,
         TIDnum,
         date,
         week,
         which(sapply(., sd, na.rm = TRUE) > 0)) %>% # only include variables that vary (i.e., sd != 0)
  as.data.frame

@@

# MiS1 <-
#   amelia(
#     dtMiS1,
#     m = 1,
#     ts = "TIDnum",
#     cs = "PID",
#     idvars =  c("TID", "date", "week"),
#     #noms = names(dtMiS1)[grepl(paste(varNamdtS1Noms, collapse = "|^"), names(dtMiS1))],
#     #parallel = "multicore",
#     #ncpus = parallel::detectCores(),
#     p2s = 2
#   )
# save(MiS1, file = "data/imputed/S1Imputed.RData")
# load("data/imputed/S1Imputed.RData")
```

@@

To be done once initial tests approved.

```{r MiS2}
# BASED ON MiS1
```

@@

To be done once initial tests approved.

```{r MiS3}
# BASED ON MiS1
```

@@

For the 3MPCA, we center (across participants but within time point) and normalize (within variable but across time points) all key variables. The between-person centering, ensures that all variations are around the mean trend (which is removed by the centering). The normalization within variable are important for ensuring equal variances across variables, which ensures equal weighting of the variables in the 3MPCA.

@@

For the Three-way ANOVA and the Three-mode PCA, the datasets need to be in a $n * mp$ matrix format, where $n$ is are the participants, $m$ are the time points, and $p$ are the variables. We do this by taking the prepared datasets, melting them into a id-variable-value table and casting the table into a wide format of variables_timepoint columns.

@@

dtPcaS123Imp <- MiMain$imputations[[1]] %>% 
  select(ID, date, week, study, any_of(varNamPcaS123)) %>%
  arrange(ID, TIDnum) %>%
  filter(TIDnum <= min(max(dtS1Red$TIDnum), max(dtS2Red$TIDnum), max(dtS3Red$TIDnum)))

dtMainWideC <- PCAnorm(data = dtPcaS123,
          pid = "ID",
          tid = "TIDnum",
          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>%
  select(
    ID,
    PID,
    TIDnum,
    study,
    ends_with("_gmc")
  ) %>%
  data.frame %>% 
  melt(
    .,
    id=c("ID", "PID", "TIDnum", "study")
  ) %>%
  mutate(
    variable = paste(gsub("_gmc", "", variable), stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
  ) %>%
  select(-TIDnum) %>%
  arrange(variable, ID) %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  replace(is.na(.), 0) %>%
  select(-ID, -PID, -study)
  #reshape(., idvar = "ID", timevar = "variable", direction = "wide") %>% View
  #pivot_wider(names_from = variable, values_from = value)

dtMainWide <- dtPcaS123 %>%
  select(
    -TID,
    -date,
    -week
  ) %>%
  data.frame %>% 
  melt(
    .,
    id=c("ID", "PID", "TIDnum", "study")
  ) %>%
  mutate(
    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
  ) %>%
  select(-TIDnum) %>%
  arrange(variable, ID) %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  replace(is.na(.), 0) %>%
  select(-ID, -PID, -study)

dtMainImpWide <- dtPcaS123Imp %>%
  select(
    -TID,
    -date,
    -week
  ) %>% 
  melt(
    .,
    id=c("ID", "TIDnum", "study")
  ) %>%
  mutate(
    variable = paste(variable, stringr::str_pad(TIDnum, width = 2, pad = 0), sep = "_")
  ) %>%
  select(-TIDnum, -study) %>%
  arrange(variable, ID) %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  #replace(is.na(.), 0) %>%
  select(-ID)

@@

We assess the percentages of explained variance for the person-, variable-, and time aspects --- which offers an indication of whether a 3MPCA is useful for the dataset. Based on the grand mean centered variables we conduct a fixed-effects three-way ANOVA of the person-, variable-, and time modes as well as their various interactions. We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the possible interdependence of the three modes.

```{r threeWayAnova, message=FALSE, warning=FALSE}
library(ThreeWay)
# threeWayAno <- threewayanova(
#   Y = dtMainWide,
#   n = nrow(dtMainWide), # Nr. ppt.s (A)
#   m = length(unique(dtPcaS123$TIDnum)), # Nr. time points (B)
#   #p = sum(!varNamPcaS123 %in% idVars) # Nr. variables (C)
#   p = length(names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) # Nr. variables (C)
# )

capture.output(
  threeWayAnoImp <- threewayanova(
    Y = dtMainImpWide,
    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
    m = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (B)
    p = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]) # Nr. variables (C)
  ), 
file='NUL')

options(knitr.kable.NA = '')
data.frame(SS = sapply(threeWayAnoImp, c)) %>%
  rownames_to_column(., var = "EffectNam") %>%
  mutate(RSqrd = SS / sum(SS) * 100) %>%
  add_row(
    EffectNam = "Total", 
    SS = sum(.$SS),
    RSqrd = NA
  ) %>%
  mutate(
    Effect = c(
      "Person",
      "Time",
      "Variable",
      "Person*Time",
      "Person*Variable",
      "Time*Variable",
      "Person*Time*Variable+error",
      "Total"
    )
  ) %>%
  select(Effect, everything(), -EffectNam) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "r", "c"),
      digits=2,
      col.names = c(
        "Effect",
        "Sum of Squares",
        "$R^2$"
      ),
      caption = "Three-way ANOVA Main Analysis Single Imputation") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )

@@

We find that the $Person*Time*Variable+error$ sums of squares are indeed the largest and speak towards the utility of a Three-Mode PCA. Interestingly, time itself has a relatively small percentage of explained variance and only becomes important in compination with person- or person- and variable-specific differences.

```{r panel-setup, include = FALSE}
xaringanExtra::use_panelset()
xaringanExtra::style_panelset_tabs(font_family = "inherit")
```

# **Plots per Item** {.tabset .tabset-fade}

@@

Please note that the items might be measured on different ranges. To avoid misinterpretation of variances and changes over time, we cluster variables with similar ranges in different plots where necessary.

@@

To illustrate the developments of the means,we plot the item means of all variables. We plot the means of the individual variables for each measurement occasion, summarizing the responses of all participants at any given time point. The displayed time grouping is based on the measurement index rather than the actual date to ensure comparability of the changes between each measurement as well as studies. We plot the means for bi-daily, daily, and weekly timescales.

### Studies 1-3 (Means) {.unlisted .unnumbered}

We beginn by assessing the imputed dataset of the main analysis, where participants from all three studies are combined, using the relevant variables that are available for all three studies.

```{r plotAllMean}
varClS123 <- dtPcaS123Imp %>%
  PCAnorm(data = .,
          pid = "ID",
          tid = "TIDnum",
          selection = names(.)[!names(.) %in% idVars]) %>%
  select(
    ends_with("_gmc")
  ) %>%
  data.frame %>%
  summarise_all(
    sd, na.rm = TRUE
  ) %>%
  t %>% 
  as.data.frame %>%
  kmeans(., centers = 2)

PCAnorm(data = dtPcaS123Imp,
          pid = "ID",
          tid = "TIDnum",
          selection = names(dtPcaS123)[!names(dtPcaS123) %in% idVars]) %>% 
  select(
    ID,
    TIDnum,
    ends_with("_gmc")
  ) %>%
  data.frame %>%
  melt(
    .,
    id=c("ID","TIDnum")
  ) %>%
  mutate(
    variable = gsub("_gmc", "", variable)
  ) %>%
  ggplot(., aes(x = variable, y = value, group = TIDnum, color = TIDnum)) +
  #geom_jitter() +
  stat_summary(fun=mean, geom="line") +
  labs(
    title = "Studies 1-3 Means over time [bi-daily]",
    y = "Grand Mean Centered Average",
    x = "Variable",
    color = "Timepoint",
    group = "Timepoint"
  ) +
  #coord_flip() +
  theme_Publication() +
  theme(
    axis.text.x = element_text(angle = 45, hjust=1),
    legend.key.size = unit(1, 'cm'),
    legend.key.height = unit(0.4, 'cm'),
  )

@@

### Study 1 (Means) {.unlisted .unnumbered}

@@

@@

### Study 2 (Means) {.unlisted .unnumbered}

@@

### Study 3 (Means) {.unlisted .unnumbered}

@@

@@

### Studies 1-3 (SDs) {.unlisted .unnumbered}
<!-- <font size="5">Studies 1-3 (SDs)</font>  -->

@@

### Study 1 (SDs) {.unlisted .unnumbered}

@@

### Study 2 (SDs) {.unlisted .unnumbered}

@@

### Study 3 (SDs) {.unlisted .unnumbered}

@@

# **References**
